An Open-Source Implementation of the CMPS Algorithm for Assessing Similarity of Bullets

### 2022-03-29 review_responds/review-responds-ju-hofmann2.md

---
output:
  pdf_document: default
  html_document: default
---

## Overview

We thank you for all the encouraging words in the reviews. Your encouragement means a lot!

## R Code and **cmpsR** package

> The coding style is not consistent through the article. For instance, in certain places, the objects have "." whereas the other places have "_" (cmps.result.KM, cmps_with_multi_scale). Please make sure that the coding style is kept consistent throughout the article.

Thank you for your advice. The coding style is now consistent both throughout the article and in the `cmpsR` package.

> In supplementary materials, all the functions are placed in a single file named 'func_collection.R'. Please make sure that only similar functions are placed in one file and the file is named such that the name reflects the type of functions within it. This will mean that the file 'func_collection.R' is split into multiple files.

`func_collection.R` is now removed from the supplementary materials since the most helper functions are included in the `cmpsR` package.

> line 70: there is a global assignment for current.max in the function get_CMPS. And it seems a bit confusing for the user who wants to access the code. I suggest that to avoid this global assignment when building a package. I think current.max = max(pos.df$cmps) is equivalent to what is written in the function.

This global assignment is now removed and is replaced by an equivalent for loop.

> Some codes are commented, such as line 50 and line 53 in cmps.R. I suggest removing them as
they are not executed, and they might confuse the user who wants to view the code.

All the irrelevant commented codes are removed.

> In script get_ccf4.R, function get_ccf5 and get_ccf4 appear to be mostly identical to each other, except one is calling the C++ code for computing cross-correlation, and the other is written in R.
It will be good to add a few more explanations on these two functions in the documentation.

`get_ccf5()` is now removed from the package. This `R` version of `get_ccf4()` is only needed for us to have a backup and is not used anywhere in the package. Now it's backed up outside the package.

> Hamby44_npeak_result_*.R - Line 17: it seems there is no package called CMPS (it might be cmpsR?)
>
> hamby44_npeak_result_generator.R : The parallelization in ~line 223 in is not working under the window environment, but the following code can make the program works by replacing the mclapply(...) with

We have corrected the package names in `Hamby44_npeak_result_*.R` and modified the parallelization so that it works on both Mac and Windows environments.

#### Note on the **cmpsR** package
\

Note that all the updates of the `cmpsR` package are only available on [github](https://github.com/willju-wangqian/cmpsR) and have not submitted to CRAN yet since we are looking forward to your further suggestions and advice and do not want to bother CRAN team for multiple times. Please make sure that you have the most updated version of the `cmpsR` package (version 0.1.2) from [github](https://github.com/willju-wangqian/cmpsR) before you execute any `R` codes that depends on it.

## Article

> Page 1 the last third line: had already defined the Land Engraved Areas (LEAs) in the last fifth line.

The second "Land Engraved Areas (LEAs)" is replaced by "LEAs".

> Page 4, get_segs(x,len=50). Is the len = 50 is the default argument of the function? How does the length of segments is determined? Will the performance of the algorithm change for different
values of len?

We added more details of the parameter `len` in the article. As discussed in the "Algorithm" section after the introduction of the steps, the default values of these parameters are given in the original CMPS paper (Chen et al., 2019) but no cross-validation has been done. The choices of these parameters will affect the final result of the algorithm, and the proposed evaluation framework based on the "Sum of Squares Ratio" could be used to evaluate the parameter choices.

> Page 8, cmps_signature_plot(...). It will be good to add legends within the plot for the black/red
solid line and the red dotted line, and the grey boxes in Figure 4a seem almost invisible.

We have added the legend for the lines presented in Figure 4b. And the grey boxes in Figure 4a is now darker than before.

> Figure 9 and Figure 10 on page 14 and page 15: Is there any overlap between KM and KNM? The counts for KM is very small. It might be good to convert the count into probability. (i.e. hist(...,
probability = TRUE))

Thank you for your suggestion. There is no overlap between KM and KNM in Figure 9 and Figure 10. And in order to make the counts more visible, we have converted the counts to XXX

## Reference

Z. Chen, W. Chu, J. A. Soons, R. M. Thompson, J. Song, and X. Zhao. Fired bullet signature correlation using the Congruent Matching Profile Segments (CMPS) method. Forensic Science International, 305: Article 109964, (10 pages), 2019. ISSN 0379-0738. doi: https://doi.org/10.1016/j.forsciint.2019.109964.

### 2022-03-29 submissions/resubmission-ju-hofmann2/ju-hofmann.Rmd
added the new revision and response to the reviews

    return(dd$diff == idx)
  } else {
    return(rep(FALSE, length = length(land1)))
  }
}
```

@@

A report published by the @nrc states that "[m]uch forensic evidence-including, for example, bite marks and firearm and toolmark identification---is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline." To overcome those criticisms and concerns, researchers have been making an effort to build databases and develop frameworks and algorithms that bring an objective and quantitative assessment into the field.
The database of @nistdb with digital 3D topographic scans from various studies by @brundage, @hamby, @Hamby:2019 and others, provides a great resource for researchers. Algorithms that can be validated and tested and generate quantitative results are developed for 2D and 3D surface texture [@song2005], tool marks [@ChumbleyL_Scott2010VoTM], and striae on Land Engraved Areas (LEAs) of bullets [@aoas, @pmid30444940, @cmps]. 
<!--Note that this is not an exhaustive list of these efforts.-->
The notion of bullet signatures is one of the results of these efforts, and we will use it to explain the CMPS algorithm and showcase the \pkg{cmpsR} implementation.
Bullet signatures can be extracted from bullet land engraved areas (LEAs), typically one for each land engraved area, and how those bullet signatures are extracted from scans will be discussed in the Background section.
Bullet signatures capture striation marks on the bullet in a numeric format and therefore serve as the foundation for algorithms. These scans played an important role in bringing objectivity into the field.
In the following sections we will: review the background of bullet signature comparisons, discuss how we followed the idea described by @cmps for the implementation of the CMPS algorithm, propose new metrics based on the CMPS score and a principled evaluation framework for algorithmic results comparison,  and present results of applying our implementation to real data.

@@

1.  **cut the comparison signature into consecutive, non-overlapping and equal-length basis segments:** The command \code{get\_segs(x, len=50)} implements this step: it takes bullet signature \code{x} in the format of a numeric vector and cuts it into consecutive, non-overlapping and equal-length segments of length \code{len}, which are referred to as "basis segments".

@@

Note that there are several parameters in the CMPS algorithm that are left to the users to decide, such as the length of basis segments \code{seg\_length} in step 1, the number of peaks \code{npeaks\_set} identified on each level in step 2, and the length of the tolerance zone \code{Tx} in both step 2 and 3.
In our implementation of the CMPS algorithm we used the parameters given in the original CMPS paper [@cmps] as the default values for these parameters.
However, the authors state that no cross-validation has been done - there might be also issues with respect to resolution of the scans.
Further research is needed, until then users are advised to think of default values as starting values and consider alternatives.
The main function that combines all steps in the CMPS algorithm described above is called \code{extract\_feature\_cmps(...)}.
Here we present it with its default parameters.

@@

p1 <- metric_plot_helper(CMPS_hamby252_results_seg$cmps.table[[2]], "cmps.max",
  subtitle = "npeaks_set: 5-3-1, seg_length: 50"
) +
  xlab(labels[1])
p2 <- metric_plot_helper(CMPS_hamby252_results_seg$cmps.table[[2]], "cmps.maxbar",
  subtitle = "npeaks_set: 5-3-1, seg_length: 50"
) +
  xlab(labels[4])

p.result1_252 <- ggarrange(
  plotlist = list(p1, p2), nrow = 1, ncol = 2,
  common.legend = TRUE, legend = "bottom"
)
p.result1_252 <- annotate_figure(p.result1_252, top = text_grob(com.title252))
p.result1_252
```

@@

```{r result1_44, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; outliers are removed in bullet signatures; \\texttt{seg\\_length = 61}, \\texttt{Tx = 30}, \\texttt{npeaks\\_set = c(5,3,1)} ", fig.height=3, warning=FALSE}
p1 <- metric_plot_helper(CMPS_hamby44_results_seg$cmps.table[[3]], "cmps.max",
  subtitle = "npeaks_set: 5-3-1, seg_length: 61"
) +
  xlab(labels[1])
p2 <- metric_plot_helper(CMPS_hamby44_results_seg$cmps.table[[3]], "cmps.maxbar",
  subtitle = "npeaks_set: 5-3-1, seg_length: 61"
) +
  xlab(labels[4])
p.result1_44 <- ggarrange(
  plotlist = list(p1, p2), nrow = 1, ncol = 2,
  common.legend = TRUE, legend = "bottom"
)
p.result1_44 <- annotate_figure(p.result1_44, top = text_grob(com.title44))
p.result1_44
```

@@

-   `hamby*_result_generator.R`: these R scripts take the processed data of Hamby set 252 and Hamby set 44, generate preliminary results used in the paper, and save the results in the `.csv` format in the folder `data-csv`.
-   `rds_generator.R`: this R script takes the generated `.csv` files and produce `CMPSpaper_results.rds`. `CMPSpaper_results.rds` is identical to `data/CMPSpaper_results.rds` and is used generate figures and other results presented in the paper.
-   `func_collection.R`: this R script contains some helper functions used by other R scripts.

>>>>

    return(dd$diff == idx)
  } else {
    return(rep(FALSE, length = length(land1)))
  }
}

library(ggrepel)

add_labels_repel <- function(plot) {
  plotting_data <- ggplot_build(plot)$data[[1]]
  plotting_data <- plotting_data %>% filter(density != 0)
  plot + geom_text_repel(data = plotting_data,
    aes(x=.data$x, y=.data$density, label=round(.data$density, 3)),
    direction = "y"
  )
}

ADD_LABELS_HIST <- TRUE
```

@@

A report published by the @nrc states that "[m]uch forensic evidence-including, for example, bite marks and firearm and toolmark identification---is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline." To overcome those criticisms and concerns, researchers have been making an effort to build databases and develop frameworks and algorithms that bring an objective and quantitative assessment into the field.
The database of @nistdb with digital 3D topographic scans from various studies by @brundage, @hamby, @Hamby:2019 and others, provides a great resource for researchers. Algorithms that can be validated and tested and generate quantitative results are developed for 2D and 3D surface texture [@song2005], tool marks [@ChumbleyL_Scott2010VoTM], and striae on Land Engraved Areas (LEAs) of bullets [@aoas, @pmid30444940, @cmps]. 
<!--Note that this is not an exhaustive list of these efforts.-->
The notion of bullet signatures is one of the results of these efforts, and we will use it to explain the CMPS algorithm and showcase the \pkg{cmpsR} implementation.
Bullet signatures can be extracted from bullet LEAs, typically one for each land engraved area, and how those bullet signatures are extracted from scans will be discussed in the Background section.
Bullet signatures capture striation marks on the bullet in a numeric format and therefore serve as the foundation for algorithms. These scans played an important role in bringing objectivity into the field.
In the following sections we will: review the background of bullet signature comparisons, discuss how we followed the idea described by @cmps for the implementation of the CMPS algorithm, propose new metrics based on the CMPS score and a principled evaluation framework for algorithmic results comparison,  and present results of applying our implementation to real data.

@@

1.  **cut the comparison signature into consecutive, non-overlapping and equal-length basis segments:** The command \code{get\_segs(x, len=50)} implements this step: it takes bullet signature \code{x} in the format of a numeric vector and cuts it into consecutive, non-overlapping and equal-length segments of length \code{len}, which are referred to as "basis segments". \wj{Note that the parameter} \code{len} \wj{determines the length of a basis segment and thus affects the total number of basis segments, which is the upper limit of the CMPS score of a comparison. The default value of} \code{len} \wj{is 50, which will result in about 25 basis segments for the example data of the package. }

@@

Note that there are several parameters in the CMPS algorithm that \wj{will affect the final results and} are left to the users to decide, such as the length of basis segments \code{seg\_length} in step 1, the number of peaks \code{npeaks\_set} identified on each level in step 2, and the length of the tolerance zone \code{Tx} in both step 2 and 3.
In our implementation of the CMPS algorithm we used the parameters given in the original CMPS paper [@cmps] as the default values for these parameters.
However, the authors state that no cross-validation has been done - there might also be issues with respect to resolution of the scans.
Further research is needed, until then users are advised to think of default values as starting values and consider alternatives. \wj{However, the evaluation framework based on the Sum of Squares Ratio introduced in the later section could be used to evaluate the choices of these parameters. }
The main function that combines all steps in the CMPS algorithm described above is called \code{extract\_feature\_cmps(...)}.
Here we present it with its default parameters.

@@

p1 <- metric_plot_helper(
  CMPS_hamby252_results_seg$cmps.table[[2]], "cmps.max",
  subtitle = "npeaks_set: 5-3-1, seg_length: 50") +
  xlab(labels[1])
p2 <- metric_plot_helper(
  CMPS_hamby252_results_seg$cmps.table[[2]], "cmps.maxbar",
  subtitle = "npeaks_set: 5-3-1, seg_length: 50") +
  xlab(labels[4])

if (ADD_LABELS_HIST) {
  p1 <- add_labels_repel(p1)
  p2 <- add_labels_repel(p2)
}

p.result1_252 <- ggarrange(
  plotlist = list(p1, p2), nrow = 1, ncol = 2,
  common.legend = TRUE, legend = "bottom"
)

p.result1_252 <- annotate_figure(p.result1_252, top = text_grob(com.title252))
p.result1_252
```

@@

```{r result1_44, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; outliers are removed in bullet signatures; \\texttt{seg\\_length = 61}, \\texttt{Tx = 30}, \\texttt{npeaks\\_set = c(5,3,1)} ", fig.height=3, warning=FALSE}
p1 <- metric_plot_helper(
  CMPS_hamby44_results_seg$cmps.table[[3]], "cmps.max",
  subtitle = "npeaks_set: 5-3-1, seg_length: 61") +
  xlab(labels[1])
p2 <- metric_plot_helper(
  CMPS_hamby44_results_seg$cmps.table[[3]], "cmps.maxbar",
  subtitle = "npeaks_set: 5-3-1, seg_length: 61") +
  xlab(labels[4])

if (ADD_LABELS_HIST) {
  p1 <- add_labels_repel(p1)
  p2 <- add_labels_repel(p2)
}

p.result1_44 <- ggarrange(
  plotlist = list(p1, p2), nrow = 1, ncol = 2,
  common.legend = TRUE, legend = "bottom"
)
p.result1_44 <- annotate_figure(p.result1_44, top = text_grob(com.title44))
p.result1_44
```

@@

-   `hamby*_result_generator.R`: these R scripts take the processed data of Hamby set 252 and Hamby set 44, generate preliminary results used in the paper, and save the results in the `.csv` format in the folder `data-csv`. Please make sure that the package versions of `bulletxtrctr` and `cmpsR` meet the requirements.
-   `rds_generator.R`: this R script takes the generated `.csv` files and produce `CMPSpaper_results.rds`. `CMPSpaper_results.rds` is identical to `data/CMPSpaper_results.rds` and is used generate figures and other results presented in the paper.

### 2022-03-29 submissions/resubmission-ju-hofmann2/ju-hofmann.tex
added the new revision and response to the reviews

A report published by the \citet{nrc} states that ``{[}m{]}uch forensic
evidence-including, for example, bite marks and firearm and toolmark
identification---is introduced in criminal trials without any meaningful
scientific validation, determination of error rates, or reliability
testing to explain the limits of the discipline.'' To overcome those
criticisms and concerns, researchers have been making an effort to build
databases and develop frameworks and algorithms that bring an objective
and quantitative assessment into the field. The database of
\citet{nistdb} with digital 3D topographic scans from various studies by
\citet{brundage}, \citet{hamby}, \citet{Hamby:2019} and others, provides
a great resource for researchers. Algorithms that can be validated and
tested and generate quantitative results are developed for 2D and 3D
surface texture \citep{song2005}, tool marks
\citep{ChumbleyL_Scott2010VoTM}, and striae on Land Engraved Areas
(LEAs) of bullets \citep[\citet{pmid30444940}, \citet{cmps}]{aoas}. The
notion of bullet signatures is one of the results of these efforts, and
we will use it to explain the CMPS algorithm and showcase the
\pkg{cmpsR} implementation. Bullet signatures can be extracted from
bullet land engraved areas (LEAs), typically one for each land engraved
area, and how those bullet signatures are extracted from scans will be
discussed in the Background section. Bullet signatures capture striation
marks on the bullet in a numeric format and therefore serve as the
foundation for algorithms. These scans played an important role in
bringing objectivity into the field. In the following sections we will:
review the background of bullet signature comparisons, discuss how we
followed the idea described by \citet{cmps} for the implementation of
the CMPS algorithm, propose new metrics based on the CMPS score and a
principled evaluation framework for algorithmic results comparison, and
present results of applying our implementation to real data.

@@

The section ``Implementation'' will go through the algorithm and show
how to use the \pkg{cmpsR} package. A further example that illustrates
the main points is also included. In the section on 'Evaluation metrics"
we introduce new CMPS metrics that summarize land-level CMPS scores and
a sum of squares ratio that can be used to evaluate algorithmic results.
The section ``Results'' presents the results of evaluating the
\pkg{cmpsR} package using Hamby set 252 and Hamby set 44. The results
from Hamby set 252 are used to verify that our implementation is, at
least qualitatively, comparable to the algorithm described in
\citet{cmps}. Results from Hamby 44 show the need for a further
investigation of the parameter choices even in the case of bullets fired
from the same barrels. The last section covers some final discussion and
conclusions.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{cut the comparison signature into consecutive, non-overlapping
  and equal-length basis segments:} The command
  \code{get\_segs(x, len=50)} implements this step: it takes bullet
  signature \code{x} in the format of a numeric vector and cuts it into
  consecutive, non-overlapping and equal-length segments of length
  \code{len}, which are referred to as ``basis segments''.
\item
  \textbf{identify candidate positions:} For each basis segment a set of
  candidate registration positions on the comparison signature is
  identified based on the segment's similarity to the reference
  signature. In a first step, the cross-correlation function of the
  segment to the reference is calculated, then a number of positions
  with high correlation values are identified as candidate positions. In
  case multiple segment lengths are considered, the length of each basis
  segment is expanded (by default it is doubled) and these two steps are
  repeated. Only when candidate positions coincide (or are similar
  enough), they are considered further. \autoref{fig:segplots} and
  \autoref{fig:seg_all} illustrate these ideas.

@@

Note that there are several parameters in the CMPS algorithm that are
left to the users to decide, such as the length of basis segments
\code{seg\_length} in step 1, the number of peaks \code{npeaks\_set}
identified on each level in step 2, and the length of the tolerance zone
\code{Tx} in both step 2 and 3. In our implementation of the CMPS
algorithm we used the parameters given in the original CMPS paper
\citep{cmps} as the default values for these parameters. However, the
authors state that no cross-validation has been done - there might be
also issues with respect to resolution of the scans. Further research is
needed, until then users are advised to think of default values as
starting values and consider alternatives. The main function that
combines all steps in the CMPS algorithm described above is called
\code{extract\_feature\_cmps(...)}. Here we present it with its default
parameters.

>>>>

A report published by the \citet{nrc} states that ``{[}m{]}uch forensic
evidence-including, for example, bite marks and firearm and toolmark
identification---is introduced in criminal trials without any meaningful
scientific validation, determination of error rates, or reliability
testing to explain the limits of the discipline.'' To overcome those
criticisms and concerns, researchers have been making an effort to build
databases and develop frameworks and algorithms that bring an objective
and quantitative assessment into the field. The database of
\citet{nistdb} with digital 3D topographic scans from various studies by
\citet{brundage}, \citet{hamby}, \citet{Hamby:2019} and others, provides
a great resource for researchers. Algorithms that can be validated and
tested and generate quantitative results are developed for 2D and 3D
surface texture \citep{song2005}, tool marks
\citep{ChumbleyL_Scott2010VoTM}, and striae on Land Engraved Areas
(LEAs) of bullets \citep[\citet{pmid30444940}, \citet{cmps}]{aoas}. The
notion of bullet signatures is one of the results of these efforts, and
we will use it to explain the CMPS algorithm and showcase the
\pkg{cmpsR} implementation. Bullet signatures can be extracted from
bullet LEAs, typically one for each land engraved area, and how those
bullet signatures are extracted from scans will be discussed in the
Background section. Bullet signatures capture striation marks on the
bullet in a numeric format and therefore serve as the foundation for
algorithms. These scans played an important role in bringing objectivity
into the field. In the following sections we will: review the background
of bullet signature comparisons, discuss how we followed the idea
described by \citet{cmps} for the implementation of the CMPS algorithm,
propose new metrics based on the CMPS score and a principled evaluation
framework for algorithmic results comparison, and present results of
applying our implementation to real data.

@@

The section ``Implementation'' will go through the algorithm and show
how to use the \pkg{cmpsR} package. A further example that illustrates
the main points is also included. In the section on 'Evaluation
metrics'' we introduce new CMPS metrics that summarize land-level CMPS
scores and a sum of squares ratio that can be used to evaluate
algorithmic results. The section ``Results'' presents the results of
evaluating the \pkg{cmpsR} package using Hamby set 252 and Hamby set 44.
The results from Hamby set 252 are used to verify that our
implementation is, at least qualitatively, comparable to the algorithm
described in \citet{cmps}. Results from Hamby 44 show the need for a
further investigation of the parameter choices even in the case of
bullets fired from the same barrels. The last section covers some final
discussion and conclusions.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{cut the comparison signature into consecutive, non-overlapping
  and equal-length basis segments:} The command
  \code{get\_segs(x, len=50)} implements this step: it takes bullet
  signature \code{x} in the format of a numeric vector and cuts it into
  consecutive, non-overlapping and equal-length segments of length
  \code{len}, which are referred to as ``basis segments''.
  \wj{Note that the parameter} \code{len}
  \wj{determines the length of a basis segment and thus affects the total number of basis segments, which is the upper limit of the CMPS score of a comparison. The default value of}
  \code{len}
  \wj{is 50, which will result in about 25 basis segments for the example data of the package. }
\item
  \textbf{identify candidate positions:} For each basis segment a set of
  candidate registration positions on the comparison signature is
  identified based on the segment's similarity to the reference
  signature. In a first step, the cross-correlation function of the
  segment to the reference is calculated, then a number of positions
  with high correlation values are identified as candidate positions. In
  case multiple segment lengths are considered, the length of each basis
  segment is expanded (by default it is doubled) and these two steps are
  repeated. Only when candidate positions coincide (or are similar
  enough), they are considered further. \autoref{fig:segplots} and
  \autoref{fig:seg_all} illustrate these ideas.

@@

Note that there are several parameters in the CMPS algorithm that
\wj{will affect the final results and} are left to the users to decide,
such as the length of basis segments \code{seg\_length} in step 1, the
number of peaks \code{npeaks\_set} identified on each level in step 2,
and the length of the tolerance zone \code{Tx} in both step 2 and 3. In
our implementation of the CMPS algorithm we used the parameters given in
the original CMPS paper \citep{cmps} as the default values for these
parameters. However, the authors state that no cross-validation has been
done - there might also be issues with respect to resolution of the
scans. Further research is needed, until then users are advised to think
of default values as starting values and consider alternatives.
\wj{However, the evaluation framework based on the Sum of Squares Ratio introduced in the later section could be used to evaluate the choices of these parameters. }
The main function that combines all steps in the CMPS algorithm
described above is called \code{extract\_feature\_cmps(...)}. Here we
present it with its default parameters.
